---
title: "Homework 3"
subtitle: "Research in Health Economics, Spring 2025"
author: "Megan Zheng"
format:
    pdf: 
        output-file: "zheng-m-hwk3-1"
        output-ext: "pdf"
        header-includes: 
            - \usepackage{float}
            - \floatplacement{table}{H}
knitr:
    opts_chunk:
        warning: false

---
```{r}
#| include: false
#| eval: true

final.data <- readRDS("../data/output/TaxBurden_Data.rds")
load("results/hwk3_workspace.Rdata")
```
```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra)
               
```

My answers in the following file blah blah blah. Check out my repository slayyyy.

\newpage
# Summarize the Data

## Question 1
### Present a bar graph showing the proportion of states with a change in their cigarette tax in each year from 1970 to 1985.

```{r} 
#| echo: false

filtered_data <- final.data %>%
  filter(between(Year,1970, 1985))

filtered_data <- filtered_data %>%
    group_by(state) %>%
    mutate(cigarette_tax_change = tax_state - lag(tax_state, order_by = Year))

proportion_data <- filtered_data %>%
     group_by(Year) %>%
     summarize(proportion_with_change = mean(!is.na(cigarette_tax_change) & cigarette_tax_change != 0))

q1 = ggplot(proportion_data, aes(x = Year, y = proportion_with_change)) +
    geom_bar(stat = "identity") +
    labs(title = "Proportion of States with Cigarette Tax Change (1970-1985)",
             x = "Year",
             y = "Proportion of States") +
    theme_minimal()

q1
```

\newpage

## Question 2
### Plot on a single graph the average tax (in 2012 dollars) on cigarettes and the average price of a pack of cigarettes from 1970 to 2018.

```{r}
#| echo: false

filtered_data_2018 <- final.data %>%
  filter(between(Year,1970, 2018))

avg_tax_price <- filtered_data_2018 %>%
  group_by(Year) %>%
  summarize(
    avg_tax = mean(tax_dollar, na.rm = TRUE),
    avg_price = mean(cost_per_pack, na.rm = TRUE))

avg_tax_price_long <- avg_tax_price %>%
  pivot_longer(cols = c(avg_tax, avg_price), 
               names_to = "Type", values_to = "Value")

q2 = ggplot(avg_tax_price_long, aes(x = Year, y = Value, color = Type)) +
  geom_line(size = 1.2) +
  labs(title = "Average Cigarette Tax and Price (1970-2019)",
       x = "Year",
       y = "Dollars (2012 Adjusted)") +
  scale_color_manual(values = c("red", "blue"),
                     labels = c("Average Tax (2012 Dollars)", "Average Price (2012 Dollars)")) +
  theme_minimal() +
  theme(legend.title = element_blank())

q2
```
\newpage

## Question 3
### Identify the 5 states with the highest increases in cigarette prices (in dollars) over the time period. Plot the average number of packs sold per capita for those states from 1970 to 2018.

```{r}
#| echo: false
price_increase <- filtered_data_2018 %>%
  group_by(state) %>%
  summarize(price_change = max(cost_per_pack, na.rm = TRUE) - min(cost_per_pack, na.rm = TRUE)) %>%
  arrange(desc(price_change))

top_5_states <- price_increase %>% slice_head(n = 5) %>% pull(state)

top_states_data <- filtered_data_2018 %>% filter(state %in% top_5_states)

avg_packs_sold <- top_states_data %>%
  group_by(Year, state) %>%
  summarize(avg_packs = mean(sales_per_capita, na.rm = TRUE), .groups = "drop")

q3 = ggplot(avg_packs_sold, aes(x = Year, y = avg_packs, color = state)) +
  geom_line(size = 1.2) +
  labs(title = "Average Packs Sold Per Capita in Top 5 States (1970-2018)",
       x = "Year",
       y = "Average Packs Sold Per Capita") +
  theme_minimal() +
  theme(legend.title = element_blank())

q3
```
\newpage

## Question 4
### Identify the 5 states with the lowest increases in cigarette prices over the time period. Plot the average number of packs sold per capita for those states from 1970 to 2018.

```{r}
#| echo: false
price_increase <- filtered_data_2018 %>%
  group_by(state) %>%
  summarize(price_change = max(cost_per_pack, na.rm = TRUE) - min(cost_per_pack, na.rm = TRUE)) %>%
  arrange(price_change) 

bottom_5_states <- price_increase %>% slice_head(n = 5) %>% pull(state)

bottom_states_data <- filtered_data_2018 %>% filter(state %in% bottom_5_states)

avg_packs_sold <- bottom_states_data %>%
  group_by(Year, state) %>%
  summarize(avg_packs = mean(sales_per_capita, na.rm = TRUE), .groups = "drop")

q4 = ggplot(avg_packs_sold, aes(x = Year, y = avg_packs, color = state)) +
  geom_line(size = 1.2) +
  labs(title = "Average Packs Sold Per Capita in 5 States with Lowest Price Increase (1970-2018)",
       x = "Year",
       y = "Average Packs Sold Per Capita") +
  theme_minimal() +
  theme(legend.title = element_blank())

q4
```
\newpage

## Question 5
### Compare the trends in sales from the 5 states with the highest price increases to those with the lowest price increases.

blah blh blah anwer

\newpage
# Estimate ATEs

## Question 6
### Focusing only on the time period from 1970 to 1990, regress log sales on log prices to estimate the price elasticity of demand over that period. Interpret your results.

```{r}
#| echo: false
data_1970_1990 <- final.data %>%
  filter(between(Year, 1970, 1990)) %>%
  drop_na(cost_per_pack, sales_per_capita)

data_1970_1990 <- data_1970_1990 %>%
  mutate(ln_sales = log(sales_per_capita),
         ln_price = log(cost_per_pack))

q6 <- feols(ln_sales ~ ln_price | state + Year, data = data_1970_1990)
summary(q6)
```
## Question 7
### Again limiting to 1970 to 1990, regress log sales on log prices using the total (federal and state) cigarette tax (in dollars) as an instrument for log prices. Interpret your results and compare your estimates to those without an instrument. Are they different? If so, why?

```{r}
#| echo: false
q7 <- feols(ln_sales ~ 1 | state + Year | ln_price ~ tax_dollar, data = data_1970_1990)
summary(q7)
```
\newpage
## Question 8
### Show the first stage and reduced-form results from the instrument.

```{r}
#| echo: false
first_stage <- feols(ln_price ~ tax_dollar | state + Year, data = data_1970_1990)
summary(first_stage)

reduced_form <- feols(ln_sales ~ tax_dollar | state + Year, data = data_1970_1990)
summary(reduced_form)
```